Postural developments modulate children's visual access to social information

### 2018-01-31 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
edits and bib

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(egg) # plot arranging
library(lme4)
library(gridExtra)
library(feather)
library(assertthat)
```
# Introduction
[^1]: These authors contributed equally to this work.

@@

Then, however, their view of the world radically changes [@adolph2007]. Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently. These motor changes have significant consequences for what children see; crawling and walking infants simply have different views of the world. For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]. In general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].

One possibility is that these motor improvements lead to developmental cascades that impact children’s emerging social, cognitive, and linguistic abilities [@iverson2010]. These postural changes also impact how children interact with their mothers: walking (vs. crawling) infants make different kinds of object-related bids for attention from their mothers and tend to hear more action directed statements (e.g., “open it”) [@karasik2014]. Further, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies. On their account, children’s ability to stand and independently locomote may change their ability to access social information (e.g., faces, gaze)  and in turn to accelerate their own learning.

@@

### Posture and Orientation Annotation
We created custom annotations to describe the child's physical posture (e.g. standing) and the orientation of the child relative to the caregiver (e.g. far away). The child's posture was categorized as being held/carried, prone (crawling or lying), sitting, or standing. The caregiver's orientation was characterized as being close, far, or behind the child (independent of distance). For the first two annotations (close/far from the child), the caregiver could either be to the front or side of the child. All annotations were made by a trained coder using the OpenSHAPA/Datavyu software [@adolph2012]. Times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations.

>>>>

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(egg) # plot arranging
library(lme4)
library(gridExtra)
library(feather)
library(assertthat)
```
# Introduction

@@

Then, however, their view of the world radically changes [@adolph2006motor]. Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently. These motor changes have significant consequences for what children see; crawling and walking infants simply have different views of the world. For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]. In general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].

These motor improvements lead to developmental cascades that impact children’s emerging social, cognitive, and linguistic abilities in various ways [@iverson2010]. For example, postural changes impact how children interact with their mothers: walking (vs. crawling) infants make different kinds of object-related bids for attention from their mothers and tend to hear more action directed statements (e.g., “open it”) [@karasik2014]. Further, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies. On their account, children’s ability to stand and independently locomote may change their ability to access social information (e.g., faces, gaze)  and in turn to accelerate their own learning.

@@

### Posture and Orientation Annotation
We created custom annotations to describe the child's physical posture (e.g. standing) and the orientation of the child relative to the caregiver (e.g. far away). The child's posture was categorized as being held/carried, prone (crawling or lying), sitting, or standing. The caregiver's orientation was characterized as being close, far, or behind the child (independent of distance). For the first two annotations (close/far from the child), the caregiver could either be to the front or side of the child. All annotations were made by a trained coder using the OpenSHAPA/Datavyu software [@adolph2012toward]. Times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations.

### 2018-01-31 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
edits and bib

Then, however, their view of the world radically changes (K. Adolph \&
Berger, 2007). Infants' motor abilities improve dramatically near the
end of the first year of life, allowing them to locomote independently.
These motor changes have significant consequences for what children see;
crawling and walking infants simply have different views of the world.
For example, during spontaneous play in a laboratory playroom, toddlers
are more likely to look at the floor while crawling than while walking
(Franchak, Kretch, Soska, \& Adolph, 2011). In general, walking infants
tend to have full visual access to their environment and the people in
it, while crawling infants do not (Kretch, Franchak, \& Adolph, 2014).

One possibility is that these motor improvements lead to developmental
cascades that impact children's emerging social, cognitive, and
linguistic abilities (Iverson, 2010). These postural changes also impact
how children interact with their mothers: walking (vs.~crawling) infants
make different kinds of object-related bids for attention from their
mothers and tend to hear more action directed statements (e.g., ``open
it'') (Karasik, Tamis-LeMonda, \& Adolph, 2014). Further, in an
observational study, Walle \& Campos (2014) found that children who were
able to walk had both higher receptive and productive vocabularies. On
their account, children's ability to stand and independently locomote
may change their ability to access social information (e.g., faces,
gaze) and in turn to accelerate their own learning.

@@

We created custom annotations to describe the child's physical posture
(e.g.~standing) and the orientation of the child relative to the
caregiver (e.g.~far away). The child's posture was categorized as being
held/carried, prone (crawling or lying), sitting, or standing. The
caregiver's orientation was characterized as being close, far, or behind
the child (independent of distance). For the first two annotations
(close/far from the child), the caregiver could either be to the front
or side of the child. All annotations were made by a trained coder using
the OpenSHAPA/Datavyu software (K. Adolph, Gilmore, Freeman, Sanderson,
\& Millman, 2012). Times when the child was out of view of the tripod
camera were marked as uncodable and were excluded from these
annotations.

@@

The OpenPose detector also provided us with the location of an agent's
wrists, which we used as a proxy for hands for two reasons. First, as we
did not capture children's entire visual field. the presence of a wrist
is likely often indicative of the presence of a hand within the field of
view. Second, hands are often occluded by objects when caregivers are
interacting with children, yet still visually accessible by the child
and part of their joint interaction.

@@

The CNN-based pose detector (OpenPose, Cao et al., 2017; Simon, Joo,
Matthews, \& Sheikh, 2017; Wei, Ramakrishna, Kanade, \& Sheikh, 2016))
provided the locations of 18 body parts (ears, nose, wrists, etc.) and
is available at
\url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
system uses a CNN for initial anatomical detection and subsequently
applies part affinity fields (PAFs) for part association, producing a
series of body part candidates. The candidates are then matched to a
single individual and finally assembled into a pose; here, we only made
use of the body parts relevant to the face and hands (nose and wrists).

@@

To evaluate face detector performance, we hand-labeled a ``gold set'' of
labeled frames. To account for the relatively rare appearance of faces
in the dataset, we hand abeled two types of samples: a sample containing
a high density of faces (half reported by MTCNN, half by OpenPose) and a
random sample from the remaining frames. Each sample was comprised of an
equal number of frames taken from each child's video. For wrist
detections, the ``gold set'' was constructed in the same manner, except
frames with a high density of wrists came only from detections made by
OpenPose. Faces were classified as present if at least half of the face
was showing; wrists were classified as present if any part of the wrist
was showing. Precision (hits / hits + false alarms), recall (hits / hits
+ misses), and F-score (harmonic mean of precision and recall) were
calculated for all detectors and are reported in Table 2.

@@

The proportion of time infants spent sitting decreased with age, and the
proportion of time infants spent standing increased with age. Both 8
month-olds and 12 month-olds spent equivalent amounts of time
lying/crawling, which was markedly decreased in the 16-month-olds, who
spent most of their time sitting or standing (see Figure
\ref{fig:posture}). We also observed changes in children's orientation
relative to their caregivers: the 8-month-olds spent more time with
their caregiver behind them supporting their sitting positions (see
Figure \ref{fig:posture}).

@@

We examined the proportion of face and hand detections across age (see
Figure \ref{fig:detByAge}). We observed a slight U-shaped function in
face detections, such that 12-month-olds appeared to have visual acces
to slightly fewer faces faces than 8 or 16-month-olds; conversely, hand
detections perhaps appeared to generally increase with age.

@@

A summary of the coefficients of a models can be found in Table 2. While
infant's posture and orientation significantly impacted the proportion
of faces and hands that infants saw, age was not a significant predictor
in either model. Thus, these results suggest that infants' visual access
to social information is modulated by their posture and orientation,
which is in turn a function of their general locomotor development.

@@

Finally, we explored how face and hand detections changed during object
labeling events as a function of infant's posture and orientation. We
analyzed a four-second window around each labeling event (e.g., ``Look
at the {[}zem{]}!''); these labeling events were hand-annotated and
synchronized with the frame-by-frame face/hand detections. We found that
infants' posture and orientations modulated their visual access to faces
and hands during labeling events; infants who were sitting or standing
were more likely to have visual access to this social information (see
Figure \ref{fig:detByNaming}). However, we did not find that infants saw
particularily more faces or hands during naming events relative to
baseline (avg. difference in proportion of hands, 8 m.o. = 0, 12 m.o. =
-0.003, 16 m.o. = -0.014; avg. difference in proportion of faces, 8 m.o.
= 0.003, 12 m.o. = 0.01, 16 m.o. = 0.021).

@@

\caption[Proportion face and hand detections around a naming instance ('Look, a Zem']{Proportion face and hand detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture. Error bars represent non-parametric bootstrapped 95 percent confidence intervals.}\label{fig:detByNaming}
\end{figure}
\end{CodeChunk}

@@

\hypertarget{refs}{}
\hypertarget{ref-adolph2007}{}
Adolph, K., \& Berger, S. (2007). Motor development. In \emph{Handbook
of child psychology}. Wiley Online Library.

\hypertarget{ref-adolph2012}{}
Adolph, K., Gilmore, R., Freeman, C., Sanderson, P., \& Millman, D.
(2012). Toward open behavioral science. \emph{Psychological Inquiry},
\emph{23}(3), 244--247.

>>>>

Then, however, their view of the world radically changes (Adolph \&
Berger, 2006). Infants' motor abilities improve dramatically near the
end of the first year of life, allowing them to locomote independently.
These motor changes have significant consequences for what children see;
crawling and walking infants simply have different views of the world.
For example, during spontaneous play in a laboratory playroom, toddlers
are more likely to look at the floor while crawling than while walking
(Franchak, Kretch, Soska, \& Adolph, 2011). In general, walking infants
tend to have full visual access to their environment and the people in
it, while crawling infants do not (Kretch, Franchak, \& Adolph, 2014).

These motor improvements lead to developmental cascades that impact
children's emerging social, cognitive, and linguistic abilities in
various ways (Iverson, 2010). For example, postural changes impact how
children interact with their mothers: walking (vs.~crawling) infants
make different kinds of object-related bids for attention from their
mothers and tend to hear more action directed statements (e.g., ``open
it'') (Karasik, Tamis-LeMonda, \& Adolph, 2014). Further, in an
observational study, Walle \& Campos (2014) found that children who were
able to walk had both higher receptive and productive vocabularies. On
their account, children's ability to stand and independently locomote
may change their ability to access social information (e.g., faces,
gaze) and in turn to accelerate their own learning.

@@

We created custom annotations to describe the child's physical posture
(e.g.~standing) and the orientation of the child relative to the
caregiver (e.g.~far away). The child's posture was categorized as being
held/carried, prone (crawling or lying), sitting, or standing. The
caregiver's orientation was characterized as being close, far, or behind
the child (independent of distance). For the first two annotations
(close/far from the child), the caregiver could either be to the front
or side of the child. All annotations were made by a trained coder using
the OpenSHAPA/Datavyu software (Adolph, Gilmore, Freeman, Sanderson, \&
Millman, 2012). Times when the child was out of view of the tripod
camera were marked as uncodable and were excluded from these
annotations.

@@

The OpenPose detector also provided us with the location of an agent's
wrists, which we used as a proxy for hands for two reasons. First, as we
did not capture children's entire visual field, the presence of a wrist
is likely often indicative of the presence of a hand within the field of
view. Second, hands are often occluded by objects when caregivers are
interacting with children, yet still visually accessible by the child
and part of their joint interaction.

@@

The CNN-based pose detector (OpenPose; Cao et al., 2017; Simon, Joo,
Matthews, \& Sheikh, 2017; Wei, Ramakrishna, Kanade, \& Sheikh, 2016))
provided the locations of 18 body parts (ears, nose, wrists, etc.) and
is available at
\url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
system uses a CNN for initial anatomical detection and subsequently
applies part affinity fields (PAFs) for part association, producing a
series of body part candidates. The candidates are then matched to a
single individual and finally assembled into a pose; here, we only made
use of the body parts relevant to the face and hands (nose and wrists).

@@

To evaluate face detector performance, we hand-labeled a ``gold set'' of
labeled frames. To account for the relatively rare appearance of faces
in the dataset, we hand-labeled two types of samples: a sample
containing a high density of faces (half reported by MTCNN, half by
OpenPose) and a random sample from the remaining frames. Each sample was
comprised of an equal number of frames taken from each child's video.
For wrist detections, the ``gold set'' was constructed in the same
manner, except frames with a high density of wrists came only from
detections made by OpenPose. Faces were classified as present if at
least half of the face was showing; wrists were classified as present if
any part of the wrist was showing. Precision (hits / hits + false
alarms), recall (hits / hits + misses), and F-score (harmonic mean of
precision and recall) were calculated for all detectors and are reported
in Table 2.

@@

The proportion of time infants spent sitting decreased with age, and the
proportion of time infants spent standing increased with age. Both
8-month-olds and 12-month-olds spent equivalent amounts of time
lying/crawling, which was markedly decreased in the 16-month-olds, who
spent most of their time sitting or standing (see Figure
\ref{fig:posture}). We also observed changes in children's orientation
relative to their caregivers: the 8-month-olds spent more time with
their caregiver behind them supporting their sitting positions (see
Figure \ref{fig:posture}).

@@

We examined the proportion of face and hand detections across age (see
Figure \ref{fig:detByAge}). We observed a slight U-shaped function in
face detections, such that 12-month-olds appeared to have visual access
to slightly fewer faces faces than 8 or 16-month-olds; conversely, hand
detections perhaps appeared to generally increase with age.

@@

A summary of the coefficients of the models can be found in Table 2.
While infant's posture and orientation significantly impacted the
proportion of faces and hands that infants saw, age was not a
significant predictor in either model. Thus, these results suggest that
infants' visual access to social information is modulated by their
posture and orientation, which is in turn a function of their general
locomotor development.

@@

Finally, we explored how face and hand detections changed during object
labeling events as a function of infants' posture and orientation. We
analyzed a four-second window around each labeling event (e.g., ``Look
at the {[}zem{]}!''); these labeling events were hand-annotated and
synchronized with the frame-by-frame face/hand detections. We found that
infants' posture and orientations modulated their visual access to faces
and hands during labeling events; infants who were sitting or standing
were more likely to have visual access to this social information (see
Figure \ref{fig:detByNaming}). However, we did not find that infants saw
particularly more faces or hands during naming events relative to
baseline (avg. difference in proportion of hands, 8 m.o. = 0, 12 m.o. =
-0.003, 16 m.o. = -0.014; avg. difference in proportion of faces, 8 m.o.
= 0.003, 12 m.o. = 0.01, 16 m.o. = 0.021).

@@

\caption[Proportion of face and hand detections around a naming instance ('Look, a Zem']{Proportion of face and hand detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture. Error bars represent non-parametric bootstrapped 95 percent confidence intervals.}\label{fig:detByNaming}
\end{figure}
\end{CodeChunk}

@@

\hypertarget{refs}{}
\hypertarget{ref-adolph2006motor}{}
Adolph, K. E., \& Berger, S. E. (2006). Motor development.
\emph{Handbook of Child Psychology}.

\hypertarget{ref-adolph2012toward}{}
Adolph, K. E., Gilmore, R. O., Freeman, C., Sanderson, P., \& Millman,
D. (2012). Toward open behavioral science. \emph{Psychological Inquiry},
\emph{23}(3), 244--247.
